An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-08-04 poly-embed-journal/poly-embed-journal.tex

\raf{TODO: update section with formal theorem statements and new result about tight embedding}

Many structured prediction settings can be thought of as making multiple predictions at once, with a loss function that jointly measures error based on the relationship between these predictions~\cite{hazan2010direct, gao2011consistency, osokin2017structured}.
In the case of $k$ binary predictions, these settings are typically formalized by taking the predictions and outcomes to be $\pm 1$ vectors, so $\R=\Y=\{-1,1\}^k$.
One then defines a joint loss function, which is often merely a function of the set of mispredictions, meaning we may write $\ell^g(r)_y = g(\{i \in [k] : r_i \neq y_i\})$ for some set function $g:2^{[k]}\to\reals$.
For example, Hamming loss is given by $g(S) = |S|$.
In an effort to provide a general convex surrogate for these settings when $g$ is a submodular function, Yu and Blaschko~\cite{yu2018lovasz} introduce the \emph{Lov\'asz hinge}, which leverages the well-known convex Lov\'asz extension of submodular functions.
While the authors provide theoretical justification and experiments, consistency of the Lov\'asz hinge is left open, which we resolve.

Rather than formally define the Lov\'asz hinge, we defer the complete analysis to the full version of the paper~\cite{finocchiaro2019embedding}, and focus here on the $k=2$ case.
For brevity, we write $g_\emptyset := g(\emptyset)$, $g_{1,2} := g(\{1,2\})$, etc.
%\raft{Jessie: I meant \emph{our} paper :-].  Leaving this here:\cite{yu2015lovaszarxiv}}
Assuming $g$ is normalized and increasing (meaning $g_{1,2} \geq \{g_1,g_2\} \geq g_\emptyset = 0$), the Lov\'asz hinge $L:\reals^k\to\reals^\Y_+$ is given by
\begin{multline}
  \label{eq:lovasz-hinge}
  L^g(u)_y = \max\Bigl\{(1-u_1y_1)_+ g_1 + (1-u_2y_2)_+ (g_{1,2}-g_1),\\[-4pt] (1-u_2y_2)_+ g_2 + (1-u_1y_1)_+ (g_{1,2}-g_2)\Bigr\}~,
\end{multline}
where $(x)_+ = \max\{x,0\}$.
We will explore the range of values of $g$ for which $L^g$ is consistent, where the link function $\psi:\reals^2\to\{-1,1\}^2$ is fixed as $\psi(u)_i = \sgn(u_i)$, with ties broken arbitrarily.

Let us consider the coefficients $g_\emptyset = 0$, $g_1 = g_2 = g_{1,2} = 1$, for which $\ell^g$ is merely 0-1 loss on $\Y$.
For consistency, for any distribution $p\in\simplex$, we must have that whenever $u \in \argmin_{u'\in\reals^2} \inprod{p}{L^g(u')}$, the outcome $\psi(u)$ must be the most likely, i.e., in $\argmax_{y\in\Y} p(y)$.
Simplifying eq.~\eqref{eq:lovasz-hinge}, however, we have
\begin{equation}
  \label{eq:lovasz-hinge-abstain}
  L^g(u)_y = \max\bigl\{(1-u_1y_1)_+,(1-u_2y_2)_+\bigr\} = \max\bigl\{1-u_1y_1,1-u_2y_2,0\bigr\}~,
\end{equation}
which is exactly the abstain surrogate~\eqref{eq:abstain-surrogate} for $d=2$.
We immediately conclude that $L^g$ cannot be consistent with $\ell^g$, as the origin will be the unique optimal report for $L^g$ under distributions with $p_y < 0.5$ for all $y$, and one can simply take a distribution which disagrees with the way ties are broken in $\psi$.
For example, if we take $\sgn(0) = 1$, then under $p((1,1)) = p((1,-1)) = p((-1,1)) = 0.2$ and $p((-1,-1)) = 0.4$, we have $\{0\} = \argmin_{u\in\reals^2} \inprod{p}{L^g(u)}$, yet we also have $\psi(0) = (1,1) \notin \{(-1,-1)\} = \argmin_{r\in\R} \inprod{p}{\ell^g(r)}$.

In fact, this example is typical: using our embedding framework, and characterizing when $0\in\reals^2$ is an embedded point, one can show that $L^g$ is consistent if and only if $g_{1,2} = g_1 + g_2$.
Moreover, in this linear case, which corresponds to $g$ being \emph{modular}, the Lov\'asz hinge reduces to weighted Hamming loss, which is trivially consistent from the consistency of hinge loss for 0-1 loss.
In the full version of the paper~\cite{finocchiaro2019embedding}, we generalize this observation for all $k$: $L^g$ is consistent if and only if $g$ is modular.
In other words, even for $k>2$, the only consistent Lov\'asz hinge is weighted Hamming loss.
These results cast doubt on the effectiveness of the Lov\'asz hinge in practice.

>>>>

Many structured prediction settings can be thought of as making multiple predictions at once, with a loss function that jointly measures error based on the relationship between these predictions~\cite{hazan2010direct, gao2011consistency, osokin2017structured}.
In the case of $k$ binary predictions, these settings are typically formalized by taking the predictions and outcomes to be $\pm 1$ vectors, so $\R=\Y=\{-1,1\}^k$.
One then defines a joint loss function, which is often merely a function of the set of mispredictions, meaning we may write $\ell^f(r)_y = f(\{i \in [k] : r_i \neq y_i\})$ for some set function $f:2^{[k]}\to\reals$.
For example, Hamming loss is given by $f(S) = |S|$.
In an effort to provide a general convex surrogate for these settings when $f$ is a submodular function, Yu and Blaschko~\cite{yu2018lovasz} introduce a surrogate $L^f$ which they dub the \emph{Lov\'asz hinge}, as it leverages the well-known convex Lov\'asz extension of submodular functions.
While the authors provide theoretical justification and experiments, consistency of the Lov\'asz hinge is left open, which we resolve.
As even the formal definition of the Lov\'asz hinge is quite involved, we defer the complete analysis to \S~\ref{app:lovasz} and here state the main results.

As a warm-up, let us introduce the form of the Lov\'asz hinge $L^f$ for the case $k=2$, as one can build intuition from this case.
For brevity, we write $f_\emptyset := f(\emptyset)$, $f_{1,2} := f(\{1,2\})$, etc.
%\raft{Jessie: I meant \emph{our} paper :-].  Leaving this here:\cite{yu2015lovaszarxiv}}
Assuming $f$ is normalized and increasing (meaning $f_{1,2} \geq \{f_1,f_2\} \geq f_\emptyset = 0$), the Lov\'asz hinge $L:\reals^k\to\reals^\Y_+$ is given by
\begin{multline}
  \label{eq:lovasz-hinge}
  L^f(u)_y = \max\Bigl\{(1-u_1y_1)_+ f_1 + (1-u_2y_2)_+ (f_{1,2}-f_1),\\[-4pt] (1-u_2y_2)_+ f_2 + (1-u_1y_1)_+ (f_{1,2}-f_2)\Bigr\}~,
\end{multline}
where $(x)_+ = \max\{x,0\}$.
The link function $\psi:\reals^2\to\{-1,1\}^2$ is fixed as $\psi(u)_i = \sgn(u_i)$, with ties broken arbitrarily.

We are interested in characterizing the functions $f$ for which $L^f$ is consistent.
As an example, consider the coefficients $f_\emptyset = 0$, $f_1 = f_2 = f_{1,2} = 1$, for which $\ell^f$ is merely 0-1 loss on $\Y$.
For consistency, for any distribution $p\in\simplex$, we must have that whenever $u \in \argmin_{u'\in\reals^2} \inprod{p}{L^f(u')}$, the outcome $\psi(u)$ must be the most likely, i.e., in $\argmax_{y\in\Y} p(y)$.
Simplifying eq.~\eqref{eq:lovasz-hinge}, however, we have
\begin{equation}
  \label{eq:lovasz-hinge-abstain}
  L^f(u)_y = \max\bigl\{(1-u_1y_1)_+,(1-u_2y_2)_+\bigr\} = \max\bigl\{1-u_1y_1,1-u_2y_2,0\bigr\}~,
\end{equation}
which is exactly the abstain surrogate~\eqref{eq:abstain-surrogate} for $d=2$.
We immediately conclude that $L^f$ cannot be consistent with $\ell^f$, as the origin will be the unique optimal report for $L^f$ under distributions with $p_y < 0.5$ for all $y$, and one can simply take a distribution which disagrees with the way ties are broken in $\psi$.
For example, if we take $\sgn(0) = 1$, then under $p((1,1)) = p((1,-1)) = p((-1,1)) = 0.2$ and $p((-1,-1)) = 0.4$, we have $\{0\} = \argmin_{u\in\reals^2} \inprod{p}{L^f(u)}$, yet we also have $\psi(0) = (1,1) \notin \{(-1,-1)\} = \argmin_{r\in\R} \inprod{p}{\ell^f(r)}$.

In fact, this example is typical: % using our embedding framework, and characterizing when $0\in\reals^2$ is an embedded point,
one can show that $L^f$ is consistent if and only if $f_{1,2} = f_1 + f_2$.
Moreover, in this linear case, which corresponds to $f$ being \emph{modular}, the Lov\'asz hinge reduces to weighted Hamming loss, which is trivially consistent from the consistency of hinge loss for 0-1 loss.
In \S~\ref{app:lovasz}, we generalize this observation for all $k$.
\begin{theorem}
	Let $f$ be submodular, normalized, and increasing.
	Then $(L^f,\psi)$ is consistent if and only if $f$ is modular.
\end{theorem}
In other words, even for $k>2$, the only consistent Lov\'asz hinge is weighted Hamming loss.
This result casts doubt on the practical effectiveness of the Lov\'asz hinge as a surrogate for $\ell^f$.

In light of this negative result, together with the fact that every polyhedral surrogate embeds some target loss (Theorem~\ref{thm:poly-embeds-discrete}), we now turn to the question: what target problem, then, \emph{is} the Lov\'asz hinge consistent for?
We show in \S~\ref{app:lovasz} that the answer is a variant of abstain loss where one can abstain on a set of labels $B\subseteq[k]$.
Recall that the symmetric difference is defined by $S\triangle T = (S\setminus T) \cup (T\setminus S)$.
\begin{theorem}
	Let $\hat\ell^f:\hat\R\times\Y\to\reals_+$, where $\hat\R = \{(A,B) \in 2^{[k]} \times 2^{[k]}: A\cap B=\emptyset\}$, be given by
	\begin{equation}
	\hat\ell^f((A,B),S) = f(A\triangle S\setminus B) + f(A\triangle S\cup B)~.
	\end{equation}
	Then the Lov\'asz hinge $L^f$ embeds $\hat\ell^f$.
  Moreover, the embedding is tight whenever $f$ is strictly submodular.
\end{theorem}
\raf{Something about how this could actually be a useful target and surrogate, provided you actually link from the surrogate to the target and not ignore $B$.}
